### 2024-12-01 paper/paper.qmd
update graphs in paper.qmd

---
title: ""
subtitle: ""
author: 
  - Yawen Tan
thanks: "Code and data are available at: [https://github.com/YawennnnnnTan/Analysis-of-Diamond-Price]"
date: today
date-format: long
abstract: ""
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(readr)
library(arrow)
library(here)
library(ggplot2)
library(patchwork)
library(ggplot2)

raw_data <- read_delim(here("data/01-raw_data/diamondprice.txt"), delim = "\t")
cleaned_data <- read_parquet(here("data/02-analysis_data/cleaned_diamondprice.parquet"))
train_data <- read_parquet(here("data/02-analysis_data/diamondprice_train_data.parquet"))
test_data <- read_parquet(here("data/02-analysis_data/diamondprice_test_data.parquet"))

model_glm <- readRDS(here("models/glm.rds"))
model_lm <- readRDS(here("models/lm.rds"))

```

# Introduction

# Data {#sec-data}

## Measurement

## Outcome Variable

在appendix里面加入一个关于所有变量的table，然后在这里引用，说明变量类型什么的
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-price
#| fig-cap: "Graphs of Diamond price"

# Graph histogram
# Calculate the mean of diamond prices
mean_price <- mean(cleaned_data$price, na.rm = TRUE)
# Create the histogram with mean line and density curve
histogram_plot <- ggplot(cleaned_data, aes(x = price)) +
  geom_histogram(binwidth = 500, fill = "lightblue", color = "black", alpha = 0.7) + 
  geom_vline(xintercept = mean_price, linetype = "dashed", color = "salmon") + # Add a dashed vertical line for the mean
  labs(
    x = "Diamond Price",
    y = "Count"
  ) +
  theme_minimal() # Use a minimal theme for clean visualization

# Graph boxplot
boxplot <- ggplot(cleaned_data, aes(x = "", y = price)) +
  geom_boxplot(fill = "pink", color = "black", alpha = 0.7, outlier.colour = "salmon") +
  labs(
    x = "",
    y = "Diamond Price"
  ) +
  theme_minimal()

# Combine 2 plots
combined_plot <- histogram_plot | boxplot
combined_plot

```

## Predictor Variables
在appendix里面加入一个关于所有变量的table，然后在这里引用，说明变量类型什么的
解释每个变量：含义是什么，包括什么值，如何排序
###  Predictor Variables
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-color
#| fig-cap: "Graphs of Predictor Variables: Color"

# Predictor Variables: Color
# Create the histogram with mean line and density curve

color_histogram_plot <- ggplot(cleaned_data, aes(x = color)) +
  geom_bar(binwidth = 3,fill = "pink", color = "black", alpha = 0.7) + 
  #scale_x_continuous(
  #  breaks = 1:8, 
  #  labels = c("D", "E", "F", "G", "H", "I", "J", "K"))+
  labs(
    title ="Distribution of Variable: Color",
    x = "Diamond Color",
    y = "Count") +
  theme_minimal() # Use a minimal theme for clean visualization

# 绘制小提琴图
violin_plot_color <- ggplot(cleaned_data, aes(x = color, y = price)) +
  geom_violin(trim = FALSE, fill = "lightblue", alpha = 0.7) +  # 添加小提琴图
  geom_boxplot(width = 0.2, outlier.color = "salmon", outlier.shape = 8, alpha = 0.7) +  # 在小提琴图中叠加箱线图
  labs(
    title = "Violin Plot of Price by Color",
    x = "Color",
    y = "Price"
  ) +
  theme_minimal()  # 使用简洁的主题

# Combine 2 plots
combined_plot <- color_histogram_plot | violin_plot_color
combined_plot
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-carat_size
#| fig-cap: "Graphs of Predictor Variable: Carat Size"

# Graph histogram
# Calculate the mean of diamond prices
mean_size <- mean(cleaned_data$carat_size, na.rm = TRUE)
carat_size_histogram_plot <- ggplot(cleaned_data, aes(x = carat_size)) +
  geom_histogram(fill = "palevioletred", color = "black", alpha = 0.7) +
  geom_vline(xintercept = mean_size, linetype = "dashed", color = "salmon") +
  labs(
    title = "Distribution of Variable: Carat Size",
    x = "Diamond Carat Size",
    y = "Count"
  ) +
  theme_minimal() # Use a minimal theme for clean visualization

plot_carat_size <- ggplot(cleaned_data, aes(x = carat_size, y = price)) +
  geom_point(color = "lightblue", alpha = 0.7) +
  geom_smooth(method = "loess", color = "black", se = FALSE) +
  labs(
    title = "Carat Size vs Price",
    x = "Carat Size",
    y = "Diamond Price"
  ) +
  theme_minimal()

# Combine 2 plots
combined_plot <- carat_size_histogram_plot | plot_carat_size
combined_plot

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-clarity
#| fig-cap: "Graphs of Predictor Variable: Clarity"

# Graph histogram
# Create the histogram with mean line
clarity_histogram_plot <- ggplot(cleaned_data, aes(x = clarity)) +
  geom_bar(bins = 20, fill = "pink", color = "black", alpha = 0.7) +
  labs(
    title = "Distribution of Variable: Clarity",
    x = "Diamond Clarity",
    y = "Count"
  ) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

boxplot_plot_clarity <- ggplot(cleaned_data, aes(x = clarity, y = price)) +
  geom_boxplot(fill = "lightblue",outlier.color = "salmon", outlier.shape = 8, alpha = 0.7) +
  labs(
    title = "Boxplot of Price by Clarity",
    x = "Cut",
    y = "Price"
  ) +
  theme_minimal()

# Combine 2 plots
combined_plot <- clarity_histogram_plot | boxplot_plot_clarity
combined_plot

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-cut
#| fig-cap: "Graphs of Predictor Variable: Cut"

# Graph histogram
# Create the histogram with mean line
cut_histogram_plot <- ggplot(cleaned_data, aes(x = cut)) +
  geom_bar(bins = 20, fill = "palevioletred", color = "black", alpha = 0.7)+
  labs(
    title = "Distribution of Variable: Cut",
    x = "Diamond Cut",
    y = "Count"
  ) +
  theme_minimal() # Use a minimal theme for clean visualization

violin_plot_cut <- ggplot(cleaned_data, aes(x = cut, y = price)) +
  geom_violin(trim = FALSE, fill = "lightblue", alpha = 0.7) + 
  geom_boxplot(width = 0.2, outlier.cut = "salmon", outlier.shape = 8, alpha = 0.7) +
  labs(
    title = "Violin Plot of Price by Cut",
    x = "Cut",
    y = "Price"
  ) +
  theme_minimal()

# Combine 2 plots
combined_plot <- cut_histogram_plot | violin_plot_cut
combined_plot

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-correlation_heat
#| fig-cap: "Correlation: Heat Graph"

# 按 cut, color, 和 clarity 聚合 carat_size 的均值
agg_data <- aggregate(carat_size ~ cut + color + clarity, data = cleaned_data, FUN = mean)

#agg_data <- aggregate(carat_size ~ cut + color + clarity, data = #cleaned_data, FUN = median)

# 绘制分面热图
ggplot(agg_data, aes(x = cut, y = color, fill = carat_size)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightblue", high = "pink") +
  facet_wrap(~ clarity) + # 按 clarity 分面
  labs(title = "Heatmap of Carat Size by Cut, Color, and Clarity",
       x = "Cut", y = "Color", fill = "Mean Carat Size") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 20, hjust = 1))+
  theme(axis.text.y = element_text(size = 6))

```

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Data preparation
包括如何处理reponse来符合gamma。详细请见appendix

## Model set-up

Define \( y_i \) as the price of the \( i \)-th diamond. Predictors include:

\begin{align} 
- \( x_{1i} \): The carat size of the diamond (in carats).
- \( x_{2i} \): The color grade of the diamond (categorical).
- \( x_{3i} \): The cut grade of the diamond (categorical).
- \( x_{4i} \): The clarity grade of the diamond (categorical).

**Model Structure**:
1. **Response Distribution**:  
   \( y_i \mid \mu_i, \phi \sim \text{Gamma}(\mu_i, \phi) \),  
   where \( \mu_i \) is the mean price and \( \phi \) is the dispersion parameter.

2. **Link Function**:  
   \( \log(\mu_i) = \beta_0 + \beta_1 x_{1i} + \sum_{j} \beta_{2j} \cdot \text{color}_{ij} + \sum_{k} \beta_{3k} \cdot \text{cut}_{ik} + \sum_{l} \beta_{4l} \cdot \text{clarity}_{il} \).

   - \( \beta_0 \): Intercept term.  
   - \( \beta_1 \): Effect of carat size.  
   - \( \beta_{2j}, \beta_{3k}, \beta_{4l} \): Effects of color, cut, and clarity, respectively.

3. **Mean Price**:  
   \( \mu_i = \exp(\beta_0 + \beta_1 x_{1i} + \sum_{j} \beta_{2j} \cdot \text{color}_{ij} + \sum_{k} \beta_{3k} \cdot \text{cut}_{ik} + \sum_{l} \beta_{4l} \cdot \text{clarity}_{il}) \).

\end{align}

## Model justification
以及为什么选择gamma，因为data section 里reponse符合分布
We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

## Model Comparision: Generalized Linear Model vs. Linear Model

## Model Evaluation
See appendix

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of Diamond Price based on Color, Cut, Clarity and Carat Size"
#| warning: false

summary(model_glm)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-prediction_actual
#| fig-cap: "Actual vs Predicted Values"

test_data$price <- scale(test_data$price, center = TRUE, scale = TRUE)
test_data$price<-exp(test_data$price)
test_data$color <- factor(test_data$color)
test_data$cut <- factor(test_data$cut)
test_data$clarity <- factor(test_data$clarity)

test_data$predicted_price <- predict(model_glm, newdata = test_data, type = "response")

ggplot(data = test_data, aes(x = 1:nrow(test_data))) +
  geom_line(aes(y = predicted_price), color = "lightblue", size = 1, linetype = "solid") + 
  geom_point(aes(y = price), color = "palevioletred", size = 2) + 
  labs(
    x = "Index",
    y = "Price"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

## Model Formula

## Example of Prediction

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Limitations
### Data Limitations
数据分布不均
### Model Limitations

## Further Considerations

\newpage

\appendix

# Appendix {-}

# Additional data details {#sec-data-details}

## Data Summary
data section
```{r}
#| echo: false
#| eval: true
#| label: tbl-cleaned data
#| tbl-cap: "First 8 rows of Cleaned data"
#| warning: false

head_cleaned_data <- head(cleaned_data,8)
knitr::kable(head_cleaned_data, format = "simple", col.names = c('Diamond Price','Carat Size','Color', 'Clarity','Cut'))

```
Variable type

## Data Cleaning
data section
## Data Preparation and Data Split
Model Section

response variable
train test to fit the mode

# Model details {#sec-model-details}

## Model Evaluation

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(model_glm) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References