### 2024-12-02 paper/paper.qmd
change the title of graphs

---
title: ""
subtitle: ""
author: 
  - Yawen Tan
thanks: "Code and data are available at: [https://github.com/YawennnnnnTan/Analysis-of-Diamond-Price]"
date: today
date-format: long
abstract: ""
format: pdf
number-sections: true
bibliography: 
  - references .bib
toc: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(readr)
library(arrow)
library(here)
library(ggplot2)
library(patchwork)
library(ggplot2)
library(ggcorrplot)
library(dplyr)

raw_data <- read_delim(here("data/01-raw_data/diamondprice.txt"), delim = "\t")
cleaned_data <- read_parquet(here("data/02-analysis_data/cleaned_diamondprice.parquet"))
train_data <- read_parquet(here("data/02-analysis_data/diamondprice_train_data.parquet"))
test_data <- read_parquet(here("data/02-analysis_data/diamondprice_test_data.parquet"))

model_glm <- readRDS(here("models/glm.rds"))
model_lm <- readRDS(here("models/lm.rds"))

```

# Introduction

# Data {#sec-data}

## Data Overview
The data set has 2690 data and 5 variables. Variables include response variable diamond price, predictor variables Diamond Color, Diamond Carat Size, Diamond Clarity, Diamond Cut.@tbl-data summarizes the range, variable type and examples of each variable.
更详细的请见:

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

```{r}
#| echo: false
#| eval: true
#| label: tbl-data
#| tbl-cap: "Summary of Diamond Dataset"
#| warning: false

# Create a data frame summarizing the variables
summary_table <- data.frame(
  Attribute = c(
    "Diamond Price (USD)", 
    "Diamond Color", 
    "Diamond Carat Size", 
    "Diamond Clarity", 
    "Diamond Cut"
  ),
  Range_or_Levels = c(
    "1000 to 10,000 (USD)", 
    "D to K (D: Worest, K: Best)", 
    "0.30 to 2.02", 
    "SI2, SI1, VS2, VS1, VVS2, VVS1, IF (Low to High)", 
    "Ideal, Good, Very Good, Excellent (Worst to Best)"
  ),
  Data_Type = c(
    "Numeric", 
    "Categorical", 
    "Numeric", 
    "Categorical", 
    "Categorical"
  ),
  Example = c(
    "5000", 
    "G", 
    "0.8", 
    "VS1", 
    "Very Good"
  )
)

# Display the table using knitr::kable
knitr::kable(
  summary_table, 
  col.names = c("Attribute", "Range or Levels", "Data Type", "Example")
)

```

## Outcome Variable

The diamond price is a continuous numerical variable, ranging from close to 1000 USD to 10,000 USD. Through the histogram of @fig-price, we can find that the price of diamonds has obvious right skewness, and with the increase of price, the number of diamonds decreases gradually. In addition, combining the histogram and box chart in @fig-price, we can find that there is no obvious abnormal value of diamond prices. Most diamond prices are concentrated in the range of 2,500 USD to 5,000 USD, with an average price of about 4,000 USD and a median of 3,500 USD, indicating that a few high-priced diamonds have raised the overall average.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-price
#| fig-cap: "Graphs of Diamond price"

# Graph histogram
# Calculate the mean of diamond prices
mean_price <- mean(cleaned_data$price, na.rm = TRUE)
# Create the histogram with mean line and density curve
histogram_plot <- ggplot(cleaned_data, aes(x = price)) +
  geom_histogram(binwidth = 500, fill = "lightblue", color = "black", alpha = 0.7) + 
  geom_vline(xintercept = mean_price, linetype = "dashed", color = "salmon") + # Add a dashed vertical line for the mean
  labs(
    x = "Diamond Price(USD)",
    y = "Count"
  ) +
  theme_minimal() # Use a minimal theme for clean visualization

# Graph boxplot
boxplot <- ggplot(cleaned_data, aes(x = "", y = price)) +
  geom_boxplot(fill = "pink", color = "black", alpha = 0.7, outlier.colour = "salmon") +
  labs(
    x = "",
    y = "Diamond Price (USD)"
  ) +
  theme_minimal()

# Combine 2 plots
combined_plot <- histogram_plot | boxplot
combined_plot

```

## Predictor Variables
在appendix里面加入一个关于所有变量的table，然后在这里引用，说明变量类型什么的
解释每个变量：含义是什么，包括什么值，如何排序

### Color
def:
distribution:
vs price:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-color
#| fig-cap: "Graphs of Predictor Variables: Color"

# Predictor Variables: Color
# Create the histogram with mean line and density curve

color_histogram_plot <- ggplot(cleaned_data, aes(x = color)) +
  geom_bar(fill = "pink", color = "black", alpha = 0.7) + 
  labs(
    title ="Distribution of Variable: Color",
    x = "Diamond Color",
    y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(size = 12, face = "bold",hjust = 0.5 ))

# 绘制小提琴图
violin_plot_color <- ggplot(cleaned_data, aes(x = color, y = price)) +
  geom_violin(trim = FALSE, fill = "lightblue", alpha = 0.7) +  # 添加小提琴图
  geom_boxplot(width = 0.2, outlier.color = "salmon", outlier.shape = 8, alpha = 0.7) +  # 在小提琴图中叠加箱线图
  labs(
    title = "Violin Plot of Price by Color",
    x = "Diamond Color",
    y = "Diamond Price (USD)"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(size = 12, face = "bold",hjust = 0.5 ))

# Combine 2 plots
combined_plot <- color_histogram_plot | violin_plot_color
combined_plot
```

### Carat Size

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-carat_size
#| fig-cap: "Graphs of Predictor Variable: Carat Size"

# Graph histogram
# Calculate the mean of diamond prices
mean_size <- mean(cleaned_data$carat_size, na.rm = TRUE)
carat_size_histogram_plot <- ggplot(cleaned_data, aes(x = carat_size)) +
  geom_histogram(fill = "palevioletred", color = "black", alpha = 0.7) +
  geom_vline(xintercept = mean_size, linetype = "dashed", color = "salmon") +
  labs(
    title = "Distribution of Variable: Carat Size",
    x = "Diamond Carat Size",
    y = "Count"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(size = 12, face = "bold",hjust = 0.5 ))

plot_carat_size <- ggplot(cleaned_data, aes(x = carat_size, y = price)) +
  geom_point(color = "lightblue", alpha = 0.7) +
  geom_smooth(method = "loess", color = "black", se = FALSE) +
  labs(
    title = "Carat Size vs Price",
    x = "Carat Size",
    y = "Diamond Price (USD)"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(size = 12, face = "bold",hjust = 0.5))

# Combine 2 plots
combined_plot <- carat_size_histogram_plot | plot_carat_size
combined_plot

```

### Clarity

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-clarity
#| fig-cap: "Graphs of Predictor Variable: Clarity"

# Graph histogram
# Create the histogram with mean line
clarity_histogram_plot <- ggplot(cleaned_data, aes(x = clarity)) +
  geom_bar(bins = 20, fill = "pink", color = "black", alpha = 0.7) +
  labs(
    title = "Distribution of Variable: Clarity",
    x = "Diamond Clarity",
    y = "Count"
  ) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(plot.title = element_text(size = 12, face = "bold",hjust = 0.5))

violin_plot_clarity <- ggplot(cleaned_data, aes(x = clarity, y = price)) +
  geom_violin(trim = FALSE, fill = "lightblue", alpha = 0.7) + 
  geom_boxplot(width = 0.2, outlier.color = "salmon", outlier.shape = 8, alpha = 0.7) +
  labs(
    title = "Violin Plot of Price by Clarity",
    x = "Clarity",
    y = "Diamond Price(USD)"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(size = 12, face = "bold",hjust = 0.5))

# Combine 2 plots
combined_plot <- clarity_histogram_plot | violin_plot_clarity
combined_plot

```

### Cut

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-cut
#| fig-cap: "Graphs of Predictor Variable: Cut"

# Graph histogram
# Create the histogram with mean line
cut_histogram_plot <- ggplot(cleaned_data, aes(x = cut)) +
  geom_bar(bins = 20, fill = "palevioletred", color = "black", alpha = 0.7)+
  labs(
    title = "Distribution of Variable: Cut",
    x = "Diamond Cut",
    y = "Count"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(size = 12, face = "bold",hjust = 0.5))

violin_plot_cut <- ggplot(cleaned_data, aes(x = cut, y = price)) +
  geom_violin(trim = FALSE, fill = "lightblue", alpha = 0.7) + 
  geom_boxplot(width = 0.2, outlier.color = "salmon", outlier.shape = 8, alpha = 0.7) +
  labs(
    title = "Violin Plot of Price by Cut",
    x = "Diamond Cut",
    y = "Diamond Price (USD)"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(size = 12, face = "bold",hjust = 0.5))

# Combine 2 plots
combined_plot <- cut_histogram_plot | violin_plot_cut
combined_plot

```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-correlation_heat
#| fig-cap: "Correlation: Heat Graph"

# Step 1: 创建一个包含独热编码的新的数据集
correlated_data <- cleaned_data %>%
  mutate(
    color = factor(color, levels = c("D", "E", "F", "G", "H", "I", "J", "K")),
    clarity = factor(clarity, levels = c("SI2", "SI1", "VS2", "VS1", "VVS2", "VVS1", "IF")),
    cut = factor(cut, levels = c("Ideal", "Good", "Very Good", "Excellent"))
  )

# 使用 model.matrix 进行独热编码

options(contrasts = c("contr.treatment", "contr.poly")) # 默认对比方法

# 强制生成所有列
one_hot_encoded <- cbind(
  correlated_data %>% select(price, carat_size), # 数值变量
  model.matrix(~ color + clarity + cut - 1, data = correlated_data, 
               contrasts.arg = list(
                 color = contrasts(correlated_data$color, contrasts = FALSE),
                 clarity = contrasts(correlated_data$clarity, contrasts = FALSE),
                 cut = contrasts(correlated_data$cut, contrasts = FALSE)
               ))
)

# Step 2: 计算相关矩阵
cor_matrix <- cor(one_hot_encoded, use = "complete.obs")

# Step 3: 自定义列和行的顺序
custom_order <- c(
  "price", 
  paste0("color", c("D", "E", "F", "G", "H", "I", "J", "K")), # Color levels
  "carat_size", 
  paste0("clarity", c("SI2", "SI1", "VS2", "VS1", "VVS2", "VVS1", "IF")), # Clarity levels
  paste0("cut", c("Ideal", "Good", "Very Good", "Excellent")) # Cut levels
)

# 确保 custom_order 中的变量名存在于 cor_matrix 中
valid_order <- intersect(custom_order, rownames(cor_matrix))

# 重新排序矩阵的行和列
cor_matrix <- cor_matrix[valid_order, valid_order]

# Step 4: 绘制相关矩阵热图
ggcorrplot(
  cor_matrix, 
  lab = FALSE,                 # 不显示相关系数的数值
  hc.order = FALSE,            # 禁用层次聚类，保持指定顺序
  type = "lower",              # 显示下三角
  lab_size = 3,                # 数值字体大小
  colors = c("lightblue", "white", "palevioletred"), # 热图颜色
  tl.cex = 8                   # 标签字体大小
)

```

## Measurement 
钻石颜色等数据是如何得出的
然后提到数据集是如何得出的

# Model

## Data preparation
包括如何处理reponse来符合gamma。详细请见appendix

## Model set-up

Define \( y_i \) as the price of the \( i \)-th diamond. Predictors include:

\begin{itemize}
    \item \( x_{1i} \): The carat size of the diamond (in carats).
    \item \( x_{2i} \): The color grade of the diamond (categorical).
    \item \( x_{3i} \): The cut grade of the diamond (categorical).
    \item \( x_{4i} \): The clarity grade of the diamond (categorical).
\end{itemize}

**Model Structure**:
1. **Response Distribution**:  
   \[
   y_i \mid \mu_i, \phi \sim \text{Gamma}(\mu_i, \phi),
   \]
   where \( \mu_i \) is the mean price and \( \phi \) is the dispersion parameter.

2. **Link Function**:  
   \begin{align}
   \log(\mu_i) &= \beta_0 + \beta_1 x_{1i} 
   + \sum_{j} \beta_{2j} \cdot \text{color}_{ij} \nonumber \\
   &\quad + \sum_{k} \beta_{3k} \cdot \text{cut}_{ik} 
   + \sum_{l} \beta_{4l} \cdot \text{clarity}_{il}.
   \end{align}
   - \( \beta_0 \): Intercept term.  
   - \( \beta_1 \): Effect of carat size.  
   - \( \beta_{2j}, \beta_{3k}, \beta_{4l} \): Effects of color, cut, and clarity, respectively.

3. **Mean Price**:  
   \begin{align}
   \mu_i &= \exp\left(\beta_0 + \beta_1 x_{1i} 
   + \sum_{j} \beta_{2j} \cdot \text{color}_{ij} \right. \nonumber \\
   &\quad \left.+ \sum_{k} \beta_{3k} \cdot \text{cut}_{ik} 
   + \sum_{l} \beta_{4l} \cdot \text{clarity}_{il}\right).
   \end{align}

## Model justification
以及为什么选择gamma，因为data section 里reponse符合分布
We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

## Model Comparision: Generalized Linear Model vs. Linear Model

## Model Evaluation
See appendix

# Results

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of Diamond Price based on Color, Cut, Clarity and Carat Size"
#| warning: false

summary(model_glm)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.pos="H"}
#| label: fig-prediction_actual
#| fig-cap: "Actual vs Predicted Values"

test_data$price <- scale(test_data$price, center = TRUE, scale = TRUE)
test_data$price<-exp(test_data$price)
test_data$color <- factor(test_data$color)
test_data$cut <- factor(test_data$cut)
test_data$clarity <- factor(test_data$clarity)

test_data$predicted_price <- predict(model_glm, newdata = test_data, type = "response")

ggplot(data = test_data, aes(x = 1:nrow(test_data))) +
  geom_line(aes(y = predicted_price), color = "lightblue", size = 1, linetype = "solid") + 
  geom_point(aes(y = price), color = "palevioletred", size = 2) + 
  labs(
    x = "Index",
    y = "Price"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

## Model Formula

## Example of Prediction

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Limitations
### Data Limitations
数据分布不均
### Model Limitations

## Further Considerations

\newpage

\appendix

# Appendix {-}

# Additional data details {#sec-data-details}

## Data Summary
data section

Variable type

## Data Cleaning
data section
## Data Preparation and Data Split
Model Section

response variable
train test to fit the mode

# Model details {#sec-model-details}

## Model Evaluation

## Posterior predictive check

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

```

## Diagnostics

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

#plot(first_model, "trace")

#plot(first_model, "rhat")
```

\newpage

# References